---
title: 'Artists of Magic: the Gathering'
author: "Nathan Werth"
date: "`r Sys.Date()`"
output: html_document
---

Do the artists of Magic: The Gathering have preferences? Mana color, creature
types, or anything of that sort?

## Exploration

```{r loadData}
library(data.table)
library(ggplot2)

theme_set(theme_minimal())

load('../data/cards.RData')
load('../data/sets.RData')

cards <- cards[sets, on = 'setCode']
```

How many artists have their been over the years?

```{r count}
length(unique(cards$artist))
```

That's a lot. Who's in the top as far as number of cards?

```{r topCount}
artist_works <- table(cards$artist)
sorted_works <- sort(artist_works, decreasing = TRUE)
head(sorted_works)
```

I actually remember some of those names. Dan Frazier is from the older sets, and
Mark Tedin sounds familiar. I started playing when Ice Age was the new thing.
Let's see when these artists were pumping out cards.

```{r timeline}
prolific_artists <- head(names(sorted_works))

ggplot(
  cards[artist %in% prolific_artists],
  aes(x = releaseDate)
  ) +
  geom_histogram(bins = 50) +
  facet_wrap(~ artist, ncol = 3)
```

Frazier was indeed a *very* common artist in the old days, and Mark Tedin also
did a lot back then. `r prolific_artists[[1]]` did the most card art, and he did
this by keeping a steadier rate of work than most of the other prolific artists.

  list(count = .N),
  keyby = c('artist', 'setCode')
]
```